---
title: "Translate Dataset from DataOne to OBIS columns"
author: "Ben Best <ben@ecoquants.com>"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: false
    code_folding: hide---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

* dataset: [Santa Barbara Channel Marine BON: Integrated fish | Environmental Data Initiative](https://portal.edirepository.org/nis/metadataviewer?packageid=edi.5.1)
* column mappings: [Vocab mappings - Google Sheets](https://docs.google.com/spreadsheets/d/1GlUWijzR-d7wM97Yv5p9VSCVdlCXSXdkot4MR0dl4M0/edit#gid=0)

## Translate Dataset

TEMP: For now, just doing first 1,000 rows.

```{r translate, message=F, warning=F}
library(tidyverse)

# SBCMBON integrated fish
url = 'https://portal.edirepository.org/nis/dataviewer?packageid=edi.5.1&entityid=91615b931c54c3aefedd0b048b22344c'

dir_cache = 'cache'
csv_cache = file.path(dir_cache, 'd.csv')
csv_temp  = 'data/temp.csv'

if (!file.exists(csv_cache)){
  d = read_csv(url)
  write_csv(d, csv_cache)
}
d = read_csv(csv_cache) # s = spec_csv(csv_cache)

# TEMP: get first 1000 rows to start
d0 = d # TODO: replace d0 with d
d = head(d, 1000)

d2 = d %>% 
  mutate(
    identificationID     = sprintf('%s_%s_%s_%s_%d', site_id, subsite_id, proj_taxon_id, transect_id, replicate_id),
    Location             = sprintf('%s_%s', site_name, subsite_name),
    decimalLatitude      = latitude,
    decimalLongitude     = longitude,
    eventDate            = date,
    eventRemarks         = data_source,
    samplingProtocol     = sprintf('%s -- %s', sample_method, sample_subtype),
    sampleSizeValue      = ifelse(is.na(height),           area,  area * height),
    sampleSizeUnit       = ifelse(is.na(height), 'square meter', 'square meter'),
    ScientificName       = taxon_name,
    taxonID              = auth_taxon_id,
    nameAccordingToID    = auth_name,
    organismQuantity     = count,
    organismQuantityType = "individual") %>%
  select(
    identificationID, 
    Location, decimalLatitude, decimalLongitude, 
    eventDate, eventRemarks,
    samplingProtocol, sampleSizeValue, sampleSizeUnit,
    ScientificName, taxonID, nameAccordingToID,
    organismQuantity, organismQuantityType)

write_csv(d2, csv_temp)

d2 %>% 
  DT::datatable()
```

## Issues

Notes:

- `site_name` is integer
- `subsite_name` has `r sum(is.na(d0$subsite_name)) %>% scales::comma()` NAs

